Comparative study of Dask and Spark for neuroimaging pipelines\\ {\footnotesize \textsuperscript{*}Note: Sub-titles are not captured in Xplore and should not be used} \thanks{Natural Sciences and Engineering Research Council}

### 2019-06-19 latex/paper.tex
Changed figfloat to subfigure. Modified the experiment section

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{times}
\usepackage{siunitx}
\usepackage{latexsym}
\usepackage{hyperref}
\usepackage{algpseudocode}
\usepackage{algorithm}
\algnewcommand\algorithmicforeach{\textbf{for each}}
\algdef{S}[FOR]{ForEach}[1]{\algorithmicforeach\ #1\ \algorithmicdo}
\usepackage[caption=false]{subfig}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}

@@

\subsection{Applications}
%% Incrementation
\subsubsection{\textbf{Incrementation}}
As our first experiment to benchmark the frameworks, we used a simple incrementation
pipeline (see Algorithm~\ref{alg:incrementation}). It reads chunks from the BigBrain
image, increments the intensity value of each voxel by 1 for a specified amount of
iterations, and write the result to an NFS as a NifTI image. Incrementing the data
reduce the caching effect. This experiment allows us to study the behavior of the
frameworks when all input are processed independently; i.e.\ each task in the graph
only depends on the previous one (see Figure~\ref{Figure fig:tg-inc}). This mimic the
behavior of analyzing multiple independent subjects in parallel.

@@

\subsection{Experiment}

\subsection{Infrastructure}
All experiments were performed on Compute Canda Cloud.

%% Material
We compare Dask (\textbf{v1.1.4}) and Spark (\textbf{v2.4.0}) based on two metrics
their scheduling and their makespan performance. All scripts used to perform our
experiments are available at
\href{https://github.com/mathdugre/paper-big-data-engines}{https://github.com/mathdugre/paper-big-data-engines}.
For our experiments, we used Compute Canada Cloud where each instance has 8 vCPUs,
30GB RAM, 10GB bandwidth and is dedicated.

%% Performance
\subsection{Experiments}
Our experiments have for goal to see how the frameworks react to different
environments and task types.

%% Incrementation
%% Node
Firstly, we vary the number of instances in our cluster environment. We test with 1,
2, 4, and 8 instances. From this, we can evaluate the level of parallelism offered by
both frameworks.

%% Iterations
Secondly, we modify the number of iterations we perform on the increment task between
the experiment. We use 1, 10 and 100 iterations. This allows the study of the
schedulers when a different amount of task is to be performed.

%% Sleep delay
Thirdly, we change the sleep delay in the increment tasks. We fix a 1, 4 and 16 and
64 seconds of sleep time for each iteration. This allows us to test the effect of
tasks with different time lengths.

%% Chunks
Finally, we split the BigBrain in a different number of chunks. We divide it in 30,
125 and 750 chunks; around 2.5GB, 0.6GB and 0.1GB, respectively. From that, we can
evaluate the framework with tasks of different size and IO requirements.

%% Baseline
Our baseline for all experiment is 10 iterations, 4 sleep delay, 125 chunks running
on 8 instances.

%% Histogram
For the histogram experiment, we vary the number of instances, as above. We also
change the number of chunks as above.

%% BIDS Example
For the BIDS example experiment, we vary the number of instances, again as above.

% Note
Note that the code implemented might not be optimal. We hope to provide a baseline
assessment of Dask and Spark.

@@

%%% INCREMENTATION %%%
%% Instances
\subsection{Experiment 1: Number of instances}
Figure~\ref{fig:inc_worker}(a) shows that the increase in performance is not
proportional to the number of workers which differ from what we would expect; it even
decrease when we reach 8 workers. This is explained by an increase in IO overhead
correlated to the number of workers (Figure~\ref{fig:inc_worker}(b)), as more worker
access the NFS simultaneously it increases the stress put on the system which results
in bottleneck, errors and slow down.

@@

\begin{figure}[!t]
    \centering
    \subfloat[Incrementation makespan]{%
      \includegraphics[clip,width=\columnwidth]{images/inc_worker.png}%
    }

    \subfloat[Incrementation total time]{%
      \includegraphics[clip,width=\columnwidth]{images/inc_idle_worker.png}%
    }
    \caption{125 chunks, 10 iterations, 4 sec.\ sleep delay}\label{fig:inc_worker}
\end{figure}

@@

\begin{figure}[!t]
    \centering
    \subfloat[Incrementation makespan]{%
      \includegraphics[clip,width=\columnwidth]{images/inc_itr.png}%
    }

    \subfloat[Incrementation total time]{%
      \includegraphics[clip,width=\columnwidth]{images/inc_idle_itr.png}%
    }
    \caption{125 chunks, 4 sec.\ sleep delay, 8 instances}\label{fig:inc_itr}
\end{figure}

@@

%%%% CHANGE NUMBER
\begin{table}[!t]
    \renewcommand{\arraystretch}{1.3}
    \caption{Distribution of the execution time in seconds for our baseline: 125
    chunks, 4 sec.\ sleep delay, 10 iterations, and 8 workers}\label{tb:inc-sleep}
    \centering
    \begin{tabular}{llllll}
    \hline
                 & Read  & Compute & Write & Overhead & Total \\ \hline
    Spark        & 34465 & 5118    & 22679 & 17076    & 79338 \\
    Dask.bag     & 36863 & 5121    & 15449 & 21388    & 78821 \\
    Dask.delayed & 32769 & 5119    & 11373 & 26199    & 75461 \\
    Dask.futures & 31891 & 5120    & 11270 & 27984    & 76265 \\ \hline
    \end{tabular}
 \end{table}

\begin{figure}[!t]
    \centering
    \subfloat[Incrementation makespan]{%
      \includegraphics[clip,width=\columnwidth]{images/inc_sleep.png}%
    }

    \subfloat[Incrementation total time]{%
      \includegraphics[clip,width=\columnwidth]{images/inc_idle_sleep.png}%
    }
    \caption{125 chunks, 10 iterations, 8 instances}
    \label{fig:inc_sleep}
\end{figure}

% Chunks
\subsection{Experiment 1: Number of chunks}
In Figure~\ref{fig:inc_chunk} Spark is not compared for 30 chunks. This is because
Spark has a 2GB limitation in the task size it can compute. Note from
Figure~\ref{fig:inc_chunk}(b) that the compute time increase when there are more
chunks. This is expected because the sleep delay is constant throughout this
experiment and there are more compute tasks when the number of chunks increases. We
also note that in Figure~\ref{fig:inc_chunk}(b) for 30 chunks the Dask bag API has a
much lower overhead time. This is because we only calculate the idle time of the used
core. Dask bag was only using one thread per block in comparison to Dask delayed and
Dask futures which was offloading the calculations on multiple threads of the same
worker. From Figure~\ref{fig:inc_chunk} we observe that when the dataset is separated
in more (smaller) chunks it saves IO time however it increases the overhead by a
similar amount; the inverse happens when we reduce the number of chunks.

\begin{figure}[!t]
    \centering
    \subfloat[Incrementation makespan]{%
      \includegraphics[clip,width=\columnwidth]{images/inc_chunk.png}%
    }

    \subfloat[Incrementation total time]{%
      \includegraphics[clip,width=\columnwidth]{images/inc_idle_chunk.png}%
    }
    \caption{10 iterations, 4 sec.\ sleep delay, 8 instances}\label{fig:inc_chunk}
\end{figure}

%% Baseline gantt chart
\subsection{Experimentation 1: Baseline timeline}
From Table~\ref{tb:inc-base} and Figure~\ref{fig:inc_gantt}, the frameworks have
similar makespan however the time spend for each type of task differs significantly.
Spark tends to spend more time than other frameworks for IO however it has a much
lower overhead. Dask delayed and futures have the lowest IO time but a much higher
overhead. Dask bag stands in the middle for both IO and overhead. Compute time is
approximately the same for all frameworks.

\begin{table}[!t]
    \renewcommand{\arraystretch}{1.3}
    \caption{Distribution of the execution time in seconds for our baseline: 125
    chunks, 4 sec.\ sleep delay, 10 iterations, and 8 workers}\label{tb:inc-base}
    \centering
    \begin{tabular}{llllll}
    \hline
                    & Read  & Compute & Write & Overhead & Total \\ \hline
    Spark        & 34465 & 5118    & 22679 & 17076    & 79338 \\
    Dask.bag     & 36863 & 5121    & 15449 & 21388    & 78821 \\
    Dask.delayed & 32769 & 5119    & 11373 & 26199    & 75461 \\
    Dask.futures & 31891 & 5120    & 11270 & 27984    & 76265 \\ \hline
    \end{tabular}
 \end{table}

\begin{figure}[!t]
    \centering
    \subfloat[Spark execution timeline]{%
      \includegraphics[clip,width=\columnwidth]{images/spark_inc_baseline_gantt.png}%
    }

    \subfloat[Dask bag execution timeline]{%
      \includegraphics[clip,width=\columnwidth]{images/Dask_bag_inc_baseline_gantt.png}%
    }

    \subfloat[Dask delayed execution timeline]{%
      \includegraphics[clip,width=\columnwidth]{images/Dask_delayed_inc_baseline_gantt.png}%
    }

    \subfloat[Dask futures execution timeline]{%
      \includegraphics[clip,width=\columnwidth]{images/Dask_futures_inc_baseline_gantt.png}%
    }
    \caption{125 chunks, 4 sec.\ sleep delay, 8 instances}
    \label{fig:inc_gantt}
\end{figure}

@@

\begin{figure}[!t]
    \centering
    \subfloat[Histogram makespan]{%
      \includegraphics[clip,width=\columnwidth]{images/histo_instance.png}%
    }

    \subfloat[Histogram total time]{%
      \includegraphics[clip,width=\columnwidth]{images/histo_idle_instances.png}%
    }
    \caption{8 instances}\label{fig:histo_wroker}
\end{figure}

@@

\begin{figure}[!t]
    \centering
    \subfloat[Histogram makespan]{%
      \includegraphics[clip,width=\columnwidth]{images/histo_splits.png}%
    }

    \subfloat[Histogram total time]{%
      \includegraphics[clip,width=\columnwidth]{images/histo_idle_splits.png}%
    }
    \caption{10 iterations, 4 sec.\ sleep delay, 8 instances}\label{fig:histo_chunk}
\end{figure}

>>>>

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{enumitem}
\usepackage{xcolor}
\usepackage{times}
\usepackage{siunitx}
\usepackage{latexsym}
\usepackage{hyperref}
\usepackage{algpseudocode}
\usepackage{algorithm}
\usepackage{caption}
\usepackage{subcaption}
\algnewcommand\algorithmicforeach{\textbf{for each}}
\algdef{S}[FOR]{ForEach}[1]{\algorithmicforeach\ #1\ \algorithmicdo}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

\begin{document}

@@

\subsection{Applications}
%% Incrementation
\subsubsection{\textbf{Incrementation}}
As our first experiment to benchmark the frameworks, we used a simple incrementation
pipeline (see Algorithm~\ref{alg:incrementation}). It reads chunks from the BigBrain
image, increments the intensity value of each voxel by 1 for a specified amount of
iterations, and write the result to an NFS as a NifTI image. Incrementing the data
reduce the caching effect. This experiment allows us to study the behavior of the
frameworks when all input are processed independently; i.e.\ each task in the graph
only depends on the previous one (see Figure~\ref{fig:tg-inc}). This mimic the
behavior of analyzing multiple independent subjects in parallel.

@@

\subsection{Experiment}
The applications are benchmarked according to up to four parameters
(Table~\ref{tab:param}):

\textbf{Number of workers:} Varying the numbers of workers helps in assessing the
scaling of the frameworks in a cluster. Also, it helps in determining the level of
parallelism of the frameworks.

\textbf{Number of chunks:} Evaluating the frameworks on different amount of chunks
allows us to understand their behavior with different task size and IO requirements.

\textbf{Number of iterations:} We change the amount of iterations for the
incrementation~\ref{alg:incrementation} to study the effect of the job size.

\textbf{Sleep delay:} We modify the sleep delay in the
incrementation~\ref{alg:incrementation} to understand the behavior of task of
different length.

The baseline for our experiments is: 8 workers, 125 chunks, 10 iterations and 4
seconds sleep delay.

\begin{table}[!b]
    \renewcommand{\arraystretch}{1.3}
    \caption{Parameters for the experiments}\label{tab:param}
    \centering
    \begin{tabular*}{\columnwidth}{llll}
    \hline
                        & Incrementation & Histogram             & BIDS Example          \\ \hline
    \# of worker        & 1, 2, 4, 8     & 1, 2, 4, 8            & 1, 2, 4, 8            \\
    \# of chunks        & 30, 125, 750   & 30, 125, 750          & 30, 125, 750          \\
    \# of iterations    & 1, 10, 100     & \multicolumn{1}{c}{-} & \multicolumn{1}{c}{-} \\
    Sleep delay {[}s{]} & 1, 4, 16, 64   & \multicolumn{1}{c}{-} & \multicolumn{1}{c}{-} \\ \hline
    \end{tabular*}
    \end{table}

\subsection{Infrastructure}
All experiments were performed on Arbutus Cloud Compute Canda. Each node has CentOS
7.5.1804 as base operating system and have the Linux kernel version
3.10.0\-862.11.6.el7.x86\_64. The nodes have an Intel Xeon Gold 6130 processor with
30GB 2666 MHz memory. The networks has a 10GB/s bandwidth. The filesystem is 2TB NFS v4.

@@

%%% INCREMENTATION %%%
%% worker
\subsection{Experiment 1: Number of instances}
Figure~\ref{fig:inc_worker}(a) shows that the increase in performance is not
proportional to the number of workers which differ from what we would expect; it even
decrease when we reach 8 workers. This is explained by an increase in IO overhead
correlated to the number of workers (Figure~\ref{fig:inc_worker}(b)), as more worker
access the NFS simultaneously it increases the stress put on the system which results
in bottleneck, errors and slow down.

@@

\begin{figure}[!t]
    \centering
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/inc_worker.png}%
        \caption{Incrementation makespan}\label{fig:inc_ms_worker}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/inc_idle_worker.png}%
        \caption{Incrementation total time}\label{fig:inc_tt_worker}
    \end{subfigure}
    \caption{125 chunks, 10 iterations, 4 sec.\ sleep delay}\label{fig:inc_worker}
\end{figure}

% Chunks
\subsection{Experiment 1: Number of chunks}
In Figure~\ref{fig:inc_chunk} Spark is not compared for 30 chunks. This is because
Spark has a 2GB limitation in the task size it can compute. Note from
Figure~\ref{fig:inc_chunk}(b) that the compute time increase when there are more
chunks. This is expected because the sleep delay is constant throughout this
experiment and there are more compute tasks when the number of chunks increases. We
also note that in Figure~\ref{fig:inc_tt_chunk} for 30 chunks the Dask bag API has a
much lower overhead time. This is because we only calculate the idle time of the used
core. Dask bag was only using one thread per block in comparison to Dask delayed and
Dask futures which was offloading the calculations on multiple threads of the same
worker. From Figure~\ref{fig:inc_chunk} we observe that when the dataset is separated
in more (smaller) chunks it saves IO time however it increases the overhead by a
similar amount; the inverse happens when we reduce the number of chunks.

\begin{figure}[!t]
    \centering
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/inc_chunk.png}%
        \caption{Incrementation makespan}\label{fig:inc_ms_chunk}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/inc_idle_chunk.png}%
        \caption{Incrementation total time}\label{fig:inc_tt_chunk}
    \end{subfigure}
    \caption{10 iterations, 4 sec.\ sleep delay, 8 instances}\label{fig:inc_chunk}
\end{figure}

@@

\begin{figure}[!t]
    \centering
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/inc_itr.png}%
        \caption{Incrementation makespan}\label{fig:inc_ms_chunk}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/inc_idle_itr.png}%
        \caption{Incrementation total time}\label{fig:inc_tt_chunk}
    \end{subfigure}
    \caption{125 chunks, 4 sec.\ sleep delay, 8 instances}\label{fig:inc_itr}
\end{figure}

@@

%%%% CHANGE NUMBER
\begin{table}[!t]
    \renewcommand{\arraystretch}{1.3}
    \caption{Distribution of the execution time in seconds for our baseline: 125
    chunks, 4 sec.\ sleep delay, 10 iterations, and 8 workers}\label{tb:inc-sleep}
    \centering
    \begin{tabular*}{\columnwidth}{llllll}
    \hline
                 & Read  & Compute & Write & Overhead & Total \\ \hline
    Spark        & 34465 & 5118    & 22679 & 17076    & 79338 \\
    Dask.bag     & 36863 & 5121    & 15449 & 21388    & 78821 \\
    Dask.delayed & 32769 & 5119    & 11373 & 26199    & 75461 \\
    Dask.futures & 31891 & 5120    & 11270 & 27984    & 76265 \\ \hline
    \end{tabular*}
 \end{table}

\begin{figure}[!t]
    \centering
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/inc_sleep.png}%
        \caption{Incrementation makespan}\label{fig:inc_ms_chunk}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/inc_idle_sleep.png}%
        \caption{Incrementation total time}\label{fig:inc_tt_chunk}
    \end{subfigure}
    \caption{125 chunks, 10 iterations, 8 instances}\label{fig:inc_sleep}
\end{figure}

%% Baseline gantt chart
\subsection{Experimentation 1: Baseline timeline}
From Table~\ref{tb:inc_base} and Figure~\ref{fig:inc_gantt}, the frameworks have
similar makespan however the time spend for each type of task differs significantly.
Spark tends to spend more time than other frameworks for IO however it has a much
lower overhead. Dask delayed and futures have the lowest IO time but a much higher
overhead. Dask bag stands in the middle for both IO and overhead. Compute time is
approximately the same for all frameworks.

\begin{table}[!t]
    \renewcommand{\arraystretch}{1.3}
    \caption{Distribution of the execution time in seconds for our baseline: 125
    chunks, 4 sec.\ sleep delay, 10 iterations, and 8 workers}\label{tb:inc_base}
    \centering
    \begin{tabular*}{\columnwidth}{llllll}
    \hline
                    & Read  & Compute & Write & Overhead & Total \\ \hline
    Spark        & 34465 & 5118    & 22679 & 17076    & 79338 \\
    Dask.bag     & 36863 & 5121    & 15449 & 21388    & 78821 \\
    Dask.delayed & 32769 & 5119    & 11373 & 26199    & 75461 \\
    Dask.futures & 31891 & 5120    & 11270 & 27984    & 76265 \\ \hline
    \end{tabular*}
 \end{table}

\begin{figure*}[!t]
    \centering
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/spark_inc_baseline_gantt.png}
        \caption{Spark execution timeline}\label{fig:inc_spark_gantt}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/Dask_bag_inc_baseline_gantt.png}%
        \caption{Dask bag execution timeline}\label{fig:inc_dask_bag_gantt}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/Dask_delayed_inc_baseline_gantt.png}%
        \caption{Dask delayed execution timeline}\label{fig:inc_dask_delayed_gantt}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/Dask_futures_inc_baseline_gantt.png}%
        \caption{Dask futures execution timeline}\label{fig:inc_dask_futures_gantt}
    \end{subfigure}
    \caption{125 chunks, 4 sec.\ sleep delay, 8 instances}
\end{figure*}

@@

\begin{figure}[!t]
    \centering
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/histo_instance.png}%
        \caption{Histogram makespan}\label{fig:histo_ms_worker}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/histo_idle_instances.png}%
        \caption{Histogram total time}\label{fig:histo_tt_worker}
    \end{subfigure}
    \caption{125 chunks}
\end{figure}

@@

\begin{figure}[!t]
    \centering
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/histo_splits.png}%
        \caption{Histogram makespan}\label{fig:histo_ms_chunk}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/histo_idle_splits.png}%
        \caption{Histogram total time}\label{fig:histo_tt_chunk}
    \end{subfigure}
    \caption{8 workers}
\end{figure}
